MSB104_assignment1

### 2023-10-09 MSB104_assignment1.qmd
Update North and South Germany 09.10.23

The first dataset we downloaded from Eurostat was **demo_r_pjanaggr3,** which contains population data from Europe.
In our case, we have selected population data from Finland, Germany and Romania.
According to @Eurostat_2021, the dataset of demography is a set of statistical numbers each country in Europe sends yearly to Eurostat by 31 December of the following calendar year.
Furthermore, the definition of "population" is, according to @Eurostat_2021, all persons having their residences in a member state, including all live births and deaths during the year when the statistical numbers are heading to Eurostat.
Moreover, the demographic data at a regional level are separated into different groups named NUTS, which are defined as nomenclature for territorial units for statistics (@Eurostat_2021).
The purpose of the NUTS function is to subdivide the territory of the European Union into 104 regions at NUTS level 1 and break it down by to 281 at NUTS level 2 and 1348 at NUTS level 3 (@Eurostat_2021).
In this paper, we have focused on the numbers of NUTS 3 level, but the dataset containing population data has been broken down by sex and broad age groups (0-14, 15-64 and 65 or more).

The following dataset we downloaded from Eurostat was the **nama_10r_3gdp dataset**.
Furthermore, this dataset is also broken down to NUTS level 3 which focus on GDP data from the sub-regions in Finland, Germany and Romania.
According to @Eurostat_2023, GDP (Gross domsetsic product) at market prices is defined as the final result of the production activity of resident producer units (@Eurostat_2023).

@@

```{r}
# Created new dataset called GDP_DE_nuts2.
# Filtered sub-regions from Germany by using str_sub to extract a substring from the Sub_regions columns containing the letters DE for Germany.
# Creating new column called id_nuts2 to save the new data for calculating Gini coefficient, for presenting the numbers of regional inequity in Germany at NUTS2 level.
# Next step is to group the numbers of NUTS2 level sorted by the the code of the region in NUTS2 level and year.
# Finally summarise the Gini coefficient numbers of each German region in NUTS2 level by multiplying 1 000 000 with GDP divided by population. 
GDP_DE_nuts2 <- merged_data %>%
  filter(
    str_sub(Sub_regions, start = 1L, end = 2L) == "DE"
  ) |> 
  mutate(
    id_nuts2 = str_sub(Sub_regions, start = 1L, end = 4L)
  ) |> 
  group_by(id_nuts2, Year) %>%
  summarise(
    GDP_per_capita_nuts2_DE = sum(1000000*GDP)/sum(Population)
    ) 
```

```{r}
# First step is to place the location of the figure caption and name it.
# Secondly, we filter GDP_DE_nuts2 to contain integer starting from 3 and ending with 4. Moreover, in this dataset we only contain vales larger or equal to 0, but less than 50.
# Thirdly, we represent the data in GPD_DE_nuts2 by mapping them in a plot using ggplot, where the x-axis represents year and y-axis represents GPD_per_capita. Furthermore the colour is determined by id_nuts2.
# Fourthly, we exclude the geom_smooth and instead use geom_line since this gives a accurate description of our data from Germany.
#| cap-location: bottom
#| fig-cap: "GDP per year."
GDP_DE_nuts2 |>
  filter(
    as.integer(str_sub(id_nuts2, start = 3L, end = 4L)) >= 0
    )  |>
  filter(
  as.integer(str_sub(id_nuts2, start = 3L, end = 4L)) < 50
  ) |> 
  ggplot(mapping = aes(
    x = Year, 
    y = GDP_per_capita_nuts2_DE, 
    colour = id_nuts2)
    ) +
  #geom_smooth(se = FALSE, snap = 0.9)
  geom_line()
```

@@

Regional inequality shows variation in regional GDP per capita, and according to @Lessmann_and_Seidel_2017, regional inequality occurs when one region experiences different assumptions than other regions.
Moreover, when the level of economic development is high, there is significant regional inequality @Lessmann_and_Seidel_2017 .
According to @williamson_1965, one crucial difference that leads to regional inequality is the access to natural resources.
Regions rich in natural resources grow faster than regions in the rest of the country, which results in inequality.
Furthermore, openness is another critical element of why regional inequality occurs.
A region with more significant access to the international trade market than other regions with limited access will gain more from international trade if international trade costs decrease @Lessmann_and_Seidel_2017

Moreover, openness can also contribute to foreign investors investing in a more attractive region, e.g., access to natural resources.
In the end, this led to an imbalance in regional equality.
Additionally, mobility can provide regional inequality through differences in transport costs from one region to another @Lessmann_and_Seidel_2017.
The differences in transport costs are caused, e.g. by fuel consumption per kilometre.

Furthermore, another significant element creating regional inequality is bureaucracy, policy and stability in a country @Lessmann_and_Seidel_2017.
Civil war affects the country's balance of regional equality and creates different assumptions.
An example is regions in a country that experience increasing aid.
Receiving assistance from foreign aid organisations creates inequality.
Lastly, @Lessmann_and_Seidel_2017 state that an area where ethnic groups dominate as a minority can result in regional inequity.

Based on the research conducted by @Lessmann_and_Seidel_2017, considering regional inequities is essential since it can influence economic outcomes.
The research shows many examples of high- and low-income countries experiencing unbalanced regional development that results in income inequality @Lessmann_and_Seidel_2017.

@@

To calculate the population-weighted Gini coefficient for GDP for each European NUTS2 level region within Germany, Finland and Romania, we used the data on the GDP and population for the different regions and associated countries.
The Gini coefficient measures income or wealth inequality within a population @IMF_n.d. In this case, we want to calculate the Gini coefficient for GDP, which is a measure of economic inequality.
We calculated the Gini coefficient using the following formula:

@@

The population-weighted Gini coefficient provides a measure of economic inequality across all our regions and associated countries, taking into account their respective populations @IMF_n.d. A Gini coefficient close to 0 indicates a more equal distribution of income or wealth among the population.
In societies with low Gini coefficients, there is relatively less income or wealth inequality, and resources are more evenly distributed among individuals or households.
These societies tend to have a larger middle class and fewer people living in poverty or extreme wealth @IMF_n.d.

A Gini coefficient close to 1 signifies a high level of income or wealth inequality.
In societies with high Gini coefficients, a significant portion of the income or wealth is concentrated among a small percentage of the population, leading to pronounced disparities.
These societies may have a large gap between the rich and the poor, with some individuals or households having substantially more resources than others @IMF_n.d.

The zero values occur as a result of regions like in Germany where the regions is actually an foreign embassy, army base or offshore energy prodution with no permanent residence.
This means they do not have any GDP per capita @Eurostat_2021.

>>>>

The first dataset we downloaded from Eurostat was **demo_r_pjanaggr3,** which contains population data from Europe.
In our case, we have selected population data from Finland, Germany and Romania.
According to @eurostat2021, the dataset of demography is a set of statistical numbers each country in Europe sends yearly to Eurostat by 31 December of the following calendar year.
Furthermore, the definition of "population" is, according to @eurostat2021, all persons having their residences in a member state, including all live births and deaths during the year when the statistical numbers are heading to Eurostat.
Moreover, the demographic data at a regional level are separated into different groups named NUTS, which are defined as nomenclature for territorial units for statistics [@eurostat2021].
The purpose of the NUTS function is to subdivide the territory of the European Union into 104 regions at NUTS level 1 and break it down by to 281 at NUTS level 2 and 1348 at NUTS level 3 [@eurostat2021].
In this paper, we have focused on the numbers of NUTS 3 level, but the dataset containing population data has been broken down by sex and broad age groups (0-14, 15-64 and 65 or more).

The following dataset we downloaded from Eurostat was the **nama_10r_3gdp dataset**.
Furthermore, this dataset is also broken down to NUTS level 3 which focus on GDP data from the sub-regions in Finland, Germany and Romania.
According to @eurostat2023, GDP (Gross domsetsic product) at market prices is defined as the final result of the production activity of resident producer units [@eurostat2023].

@@

```{r}
# Created new dataset called GDP_DE_nuts2.
# Filtered sub-regions from Germany by using str_sub to extract a substring from the Sub_regions columns containing the letters DE for Germany.
# Creating new column called id_nuts2 to save the new data for calculating Gini coefficient, for presenting the numbers of regional inequity in Germany at NUTS2 level.
# Next step is to group the numbers of NUTS2 level sorted by the the code of the region in NUTS2 level and year.
# Finally summarise the Gini coefficient numbers of each German region in NUTS2 level by multiplying 1 000 000 with GDP divided by population. 
GDP_DE_nuts2 <- merged_data %>%
  filter(
    str_sub(Sub_regions, start = 1L, end = 2L) == "DE"
  ) |> 
  mutate(
    id_nuts2 = fct(str_sub(Sub_regions, start = 1L, end = 4L))
  ) |> 
  group_by(id_nuts2, Year) %>%
  summarise(
    GDP_per_capita_nuts2_DE = sum(1000000*GDP)/sum(Population)
    ) 
```

```{r}
# Since Germany contains many regions, we have deceided to present them in two new values.
# DEz1North and DEz2SOUTH are the two values where we have devided Germany in South and North. 
DEz1NORTH <- c("DE30", "DE40", "DE50", "DE60", "DE80", "DE91", "DE92", "DE93", "DE94", "DEA1", "DEA2", "DEA3", "DEA4", "DEA5", "DEE0", "DEF0", "DEG0") 
DEz2SOUTH <- c("DE11", "DE12", "DE13", "DE14", "DE21", "DE22", "DE23", "DE24", "DE25", "DE26", "DE27", "DE71", "DE72", "DE73", "DEB1", "DEB2", "DEB3", "DEC0", "DED2", "DED4", "DED5")
```

```{r}
# First step is to place the location of the figure caption and name it.
# Secondly, we filter GDP_DE_nuts2 to contain integer starting from 3 and ending with 4. Moreover, in this dataset we only contain vales larger or equal to 0, but less than 50.
# Thirdly, we represent the data in GPD_DE_nuts2 by mapping them in a plot using ggplot, where the x-axis represents year and y-axis represents GPD_per_capita. Furthermore the colour is determined by id_nuts2.
# Fourthly, we exclude the geom_smooth and instead use geom_line since this gives a accurate description of our data from Germany.
#| cap-location: bottom
#| fig-cap: "GDP per year."
GDP_DE_nuts2 |>
  filter(
    id_nuts2 %in% DEz1
  ) |> 
  ggplot(mapping = aes(
    x = Year, 
    y = GDP_per_capita_nuts2_DE, 
    colour = id_nuts2)
    ) +
  #geom_smooth(se = FALSE, snap = 0.9)
  geom_line()
```

```{r}
GDP_DE_nuts2 |>
  filter(
    id_nuts2 %in% DEz2
  ) |> 
  ggplot(mapping = aes(
    x = Year, 
    y = GDP_per_capita_nuts2_DE, 
    colour = id_nuts2)
    ) +
  #geom_smooth(se = FALSE, snap = 0.9)
  geom_line()
```

@@

Regional inequality shows variation in regional GDP per capita, and according to @lessmann_and_seidel2017, regional inequality occurs when one region experiences different assumptions than other regions.
Moreover, when the level of economic development is high, there is significant regional inequality [@lessmann_and_seidel2017].
According to @williamson1965, one crucial difference that leads to regional inequality is the access to natural resources.
Regions rich in natural resources grow faster than regions in the rest of the country, which results in inequality.
Furthermore, openness is another critical element of why regional inequality occurs.
A region with more significant access to the international trade market than other regions with limited access will gain more from international trade if international trade costs decrease [@lessmann_and_seidel2017].

Moreover, openness can also contribute to foreign investors investing in a more attractive region, e.g., access to natural resources.
In the end, this led to an imbalance in regional equality.
Additionally, mobility can provide regional inequality through differences in transport costs from one region to another [@lessmann_and_seidel2017].
The differences in transport costs are caused, e.g. by fuel consumption per kilometre.

Furthermore, another significant element creating regional inequality is bureaucracy, policy and stability in a country [@lessmann_and_seidel2017].
Civil war affects the country's balance of regional equality and creates different assumptions.
An example is regions in a country that experience increasing aid.
Receiving assistance from foreign aid organisations creates inequality.
Lastly, @lessmann_and_seidel2017 state that an area where ethnic groups dominate as a minority can result in regional inequity.

Based on the research conducted by , considering regional inequities is essential since it can influence economic outcomes.
The research shows many examples of high- and low-income countries experiencing unbalanced regional development that results in income inequality [@lessmann_and_seidel2017].

@@

To calculate the population-weighted Gini coefficient for GDP for each European NUTS2 level region within Germany, Finland and Romania, we used the data on the GDP and population for the different regions and associated countries.
The Gini coefficient measures income or wealth inequality within a population [@imf_n_d].
In this case, we want to calculate the Gini coefficient for GDP, which is a measure of economic inequality.
We calculated the Gini coefficient using the following formula:

@@

The population-weighted Gini coefficient provides a measure of economic inequality across all our regions and associated countries, taking into account their respective populations [@imfnd]. A Gini coefficient close to 0 indicates a more equal distribution of income or wealth among the population. In societies with low Gini coefficients, there is relatively less income or wealth inequality, and resources are more evenly distributed among individuals or households. These societies tend to have a larger middle class and fewer people living in poverty or extreme wealth [@imfnd]

A Gini coefficient close to 1 signifies a high level of income or wealth inequality.
In societies with high Gini coefficients, a significant portion of the income or wealth is concentrated among a small percentage of the population, leading to pronounced disparities.
These societies may have a large gap between the rich and the poor, with some individuals or households having substantially more resources than others [@imfnd].

The zero values occur as a result of regions like in Germany where the regions is actually an foreign embassy, army base or offshore energy prodution with no permanent residence.
This means they do not have any GDP per capita [@eurostat2021].
